---
title: "NLSY79 longitudinal project"
author: "Cristian T"
date: last-modified
format: 
   html:
     df-print: paged
     embed-resources: true
editor: 
  markdown: 
    wrap: 72
---

```{r load the data, warning=F, message=F}
suppressMessages(library(tidyverse))
suppressMessages(library(tibble))
suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))
suppressMessages(library(scales))
suppressMessages(library(RColorBrewer))

# Read the RData files into a data frame
load("physical_data_nlsy79.RData")
load("education_data_nlsy79.RData")
load("income_data_nlsy79.RData")
```

## Join Data

Our analysis examines income, education, and physical characteristics in
2014, so we joined the relevant datasets and restricted the year.

```{r}
data <- inner_join(income_data_nlsy79, education_data_nlsy79, 
                  by=c("CASEID", "year"))
data <- inner_join(data, physical_data_nlsy79, 
                  by=c("CASEID", "year")) %>%
  select(CASEID, year, income, education, hair, sex)
data <- data %>% filter(year==2014)
glimpse(data)
```

After joining and limiting the year to 2014, the filtered data has 12686
rows and six columns.

## Exploration of the Income Variable

We will begin by running some basic summary statistics to provide us
with a basic understanding of the distribution and characteristics of
the income data in the dataset. We first run a check for missing values
to assess data completeness. We further explore the measures of central
tendency (mean, median), dispersion (standard deviation), and range
(minimum and maximum values).

```{r exploring income}
# Check for missing values in the income column
missing_incomes <- sum(is.na(data$income))
# Mean
mean_income <- mean(data$income, na.rm = TRUE)
# Median
median_income <- median(data$income, na.rm = TRUE)
# Standard deviation
sd_income <- sd(data$income, na.rm = TRUE)
# Minimum
min_income <-min(data$income, na.rm = TRUE)
# Maximum
max_income <- max_income <- max(data$income, na.rm = TRUE)
# Printing the results
cat("\nMissing Incomes:", missing_incomes, "\n")
cat("\nMean Income:", mean_income, "\n")
cat("Median Income:", median_income, "\n")
cat("Standard Deviation of Income:", sd_income, "\n")
cat("Minimum Income:", min_income, "\n")
cat("Maximum Income:", max_income, "\n")
#overview of Summary statistics
summary(income_data_nlsy79$income)
```

Visually inspecting income we see the following plot.

```{r warning=F, message=F}
p <- ggplot(data, aes(x = income)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "black") +  # Specify fill and border color
  labs(x = "Income", y = "Frequency") +
  scale_x_continuous(labels = scales::comma) +
  theme_minimal()

ggsave("income.png", plot = p)
p
```

Based on the above graph, over 2000 responses show zero income in 2014,
but we will keep our zero values. We want to explore the level of
education and how it correlates with income.

Remove any rows with missing values in income.

```{r}
data <- data %>%
  filter(!is.na(income))
```

## Study of Education

```{r}
p <- ggplot(data = data, aes(x = education)) + geom_bar(fill = "steelblue", color = "black")
ggsave("education.png", plot = p)
p
```

This plot reveals the spurious value 95, meaning some ungraded
education. We'll exclude these from the analysis. Evidently, there are
few of these values in the education data.

```{r}
p <- ggplot(
  data = filter(data, education < 95),
  aes(x = education)
) + geom_bar(fill = "steelblue", color = "black")
ggsave("education.png", plot = p)
p
```

```{r}
data <- data %>% filter(education<95)
```

```{r}
sum(is.na(data$education))
```

There is no missing value in education now.

## Study of Gender

Let's plot sex to examine the distribution.

```{r}
p <- ggplot(data = data, aes(x = sex)) + geom_bar(fill = "steelblue", color = "black")
ggsave("sex.png", plot = p)
p
```

```{r}
sum(is.na(data$sex))
```

There is no missing value in sex now.

## Study of Hair

```{r}
p <- ggplot(data = data,  aes(x = hair) ) + geom_bar(fill = "steelblue", color = "black")
ggsave("hair.png", plot = p)
p
```

This graph shows that over 2,500 respondents have brown hair, and over
2,000 respondents have black hair.

```{r}
sum(is.na(data$hair))
```

There are 191 missing values in hair.

## Explore the Effects of Education on Income

In this section, we are looking at the connection between education
levels and income. We are trying to understand income distribution
dynamics concerning different educational milestones and the
individuals' mean income.

```{r compare income to education }
#plot showing distribution education and income
p <- ggplot(data, aes(x = as.factor(education), y = income)) +
  geom_boxplot(fill = "steelblue", color = "black") +
  labs(x = "Education (Years)", y = "Income") +
  scale_y_continuous(labels = scales::comma)
ggsave("income_education.png", plot = p)
p
```

In the above graph, we are exploring the relationship between education
levels and income. The chart presents a box plot showing income
distribution across various education levels, using blue boxes to
highlight the interquartile range of incomes within each group. The plot
suggests that there is a positive association between income and years
of education.

To further clean things up, we will exclude any education under the 8th
grade. We are performing this cleanse to comply with inaccuracies and
following the law that most states have some compulsory education laws
and adjusting for potential age allowances that some states will allow
one to drop out of school.

```{r}
p <- ggplot(
  filter(data, education >= 8  & income <= 200000),
  aes(x = as.factor(education), y = income)) +
  geom_boxplot(fill = "steelblue", color = "black") +
  labs(x = "Education (Years)", y = "Income") +
  scale_y_continuous(labels = scales::comma)
ggsave("income_education_yrs.png", plot = p)
p
```

The positive linear relationship becomes more evident after removing the
respondents with less than eight years of education, highlighting the
influence of education level on income.

```{r bar plot}
income_by_education_mean <- data %>%
  filter(income <= 200000, education >= 8) %>% 
  group_by(education) %>%
  summarize(mean_income = mean(income, na.rm = TRUE))

p <- ggplot(income_by_education_mean, 
       aes(x = as.factor(education), y = mean_income)) +
  geom_bar(stat = "identity", fill = "steelblue", color = "black") +
  labs(x = "Education (Years)", y = "Mean Income", title = "Mean Income by Education Level")
ggsave("mean_income_by_years.png", plot = p)
p
```

```{r}
income_by_education_median <- data %>%
  filter(income <= 200000, education >= 8) %>% 
  group_by(education) %>%
  summarize(median_income = median(income, na.rm = TRUE))

p <- ggplot(income_by_education_median, aes(x = as.factor(education), y = median_income)) +
  geom_bar(stat = "identity", fill = "steelblue", color = "black") +
  labs(x = "Education (Years)", y = "Median Income", title = "Median Income by Education Level")
ggsave("mean_income_by_level.png", plot = p)
p
```

Continuing our examination, we look at the relationship between
education levels and income by summarizing individuals' mean and median
income better to understand the income distribution within each
education group. The bar chart illustrates how mean and median income
differs across these academic years. Looking at both of the graphs, we
see how much those exceptionally high incomes have affected the overall
mean income, while the median incomes appears to be reflective of the
majority of individuals.

```{r simple plot to compare}
# Filter the data for the specific education levels
education_levels <- c(12, 14, 16, 18, 20)
income_by_education <- data %>%
  filter(education %in% education_levels)

# Create a bar plot for mean
p <- ggplot(income_by_education, aes(x = factor(education), y = income)) +
  geom_bar(stat = "summary", fun = "mean", fill = "steelblue", color = "black") +
  labs(x = "Education Level", y = "Mean Income") +
  scale_x_discrete(labels = c("12" = "High School", "14" = "Associates", "16" = "Bachelors", "18" = "Masters", "20" = "Doctorate"))
ggsave("mean_income_by_education.png", plot = p)
p

# Create a bar plot for median
q <- ggplot(income_by_education, aes(x = factor(education), y = income)) +
  geom_bar(stat = "summary", fun = "median", fill = "lightblue", color = "black") +
  labs(x = "Education Level", y = "Median Income") +
  scale_x_discrete(labels = c("12" = "High School", "14" = "Associates", "16" = "Bachelors", "18" = "Masters", "20" = "Doctorate"))
ggsave("mean_income_by_education.png", plot = q)
q
```

This second series of graphs focuses on specific education levels (12,
14, 16, 18, 20) corresponding to milestones such as high school,
associate's degrees, bachelor's degrees, MBAs, and doctorates. We
calculate and visualize the mean and median income for individuals with
these distinct education levels.

```{r add additional insights}

# Calculate summary statistics for income by education level
summary_stats <- income_by_education %>%
  group_by(education) %>%
  summarise(mean_income = mean(income, na.rm = TRUE),
            sd_income = sd(income, na.rm = TRUE),
            n = n())

# Create a bar plot with error bars
p <- ggplot(summary_stats, aes(x = factor(education), y = mean_income)) +
  geom_bar(stat = "identity", fill = "steelblue", color = "black", width = 0.5) +
  geom_errorbar(aes(ymin = mean_income - sd_income / sqrt(n), ymax = mean_income + sd_income / sqrt(n)),
                width = 0.2, color = "red") +
  labs(x = "Education Level", y = "Mean Income", title = "Mean Income by Education Level") +
  scale_x_discrete(labels = c("12" = "High School", "14" = "Associates", "16" = "Bachelors", "18" = "Masters", "20" = "Doctorate")) +
  theme_minimal()
ggsave("mean_income_by_education_level.png", plot = p)
p
```

We are looking at the mean income for each education level achieved in
2014. This graph shows the average income for individuals with each
corresponding education level. We added the red error bars to provide
insights into the reliability of these income estimates, showing us the
variability in income within each education level and representing the
uncertainty or variability in the average income estimates for each
education level. The height of these bars reflects the degree of
variability within each education level's income, with taller bars
signifying greater variability and shorter bars indicating less
uncertainty in the income estimate

## Comparing Income to Physical Attributes

We will begin exploring the relationship between income and physical
attributes, focusing on hair color.

```{r}
# Group hair colors into broader categories
data <- data %>%
    filter(!is.na(hair), !is.na(income)) %>%
  mutate(broad_hair_color = case_when(
    hair %in% c("blond", "light blond") ~ "blond",
    hair %in% c("brown", "light brown") ~ "brown",
    hair %in% c("grey", "other_low_count_category1", "other_low_count_category2") ~ "other",
    TRUE ~ as.character(hair)  # Keep other categories as is
  ))

# we want to see a new count of our distribution of hair color, count the number of observations for hair color
hair_color_counts_broad <- data %>%
  group_by(broad_hair_color) %>%
  summarize(observations = n())

#display counts
hair_color_counts_broad

#our observations are now updated but now we see black and brown consisting of the majority and other is 3 not sure if this is better but going with it
```

```{r}
# Calculate the mean income for each broad hair color category
mean_income_by_color <- data %>%
  group_by(broad_hair_color) %>%
  summarize(mean_income = mean(income, na.rm = TRUE))  %>%
  mutate(broad_hair_color = reorder(broad_hair_color, -mean_income))

# Define custom colors for each hair color category
hair_color_colors <- c(
  "other" = "burlywood", 
  "blond" = "darkgoldenrod",
  "brown" = "brown",
  "black" = "black",
  "red" = "red"
)

# Create a bar plot for mean income by broad hair color
p <- ggplot(mean_income_by_color, 
       aes(x = broad_hair_color, y = mean_income, fill = broad_hair_color)) +
  geom_bar(stat = "identity") +
  labs(title = "Mean Income by Broad Hair Color Categories", x = "Broad Hair Color Category", y = "Mean Income") +
  scale_fill_manual(values = hair_color_colors) +  # Use custom colors
  theme_minimal() + labs(fill = "Hair Color")  # Rename the legend

ggsave("mean_income_by_color.png", plot = p)
p
```

We examine the distribution of hair colors in the dataset and find that
the hair colors appear imbalanced. Black and brown have the most
observations, but we are limited in variety for the colors red, grey,
blond, and light blond. These results will ultimately create an
imbalance and affect the statistical significance and reliability when
analyzing the relationship between hair color and income.

## Mean Income by Broad Hair Color Categories

```{r make some adjustments}

#the other category looks misleading so we should filter this out
# Calculate the mean income for each broad hair color category, excluding "other"
mean_income_by_color_exclude_other <- data %>%
  filter(broad_hair_color != "other") %>%
  group_by(broad_hair_color) %>%
  summarize(mean_income = mean(income, na.rm = TRUE))

# Reorder the broad hair color categories by mean income in descending order
mean_income_by_color_exclude_other <- mean_income_by_color_exclude_other %>%
  mutate(broad_hair_color = reorder(broad_hair_color, -mean_income))

# Create a bar plot for mean income by broad hair color with custom colors
p <- ggplot(mean_income_by_color_exclude_other, aes(x = broad_hair_color, y = mean_income, fill = broad_hair_color)) +
  geom_bar(stat = "identity") +
  labs(title = "Mean Income by Hair Color Categories (Excluding 'Other')", x = "Hair Color Category", y = "Mean Income") +
  scale_fill_manual(values = hair_color_colors) +  # Use custom colors
 theme_minimal() + labs(fill = "Hair Color")  # Rename the legend

ggsave("mean_income_by_color_exclude_other.png", plot = p)
p


# Find the median income for the black hair color
reference_median <- mean_income_by_color_exclude_other %>%
  filter(broad_hair_color == "black") %>%
  pull(mean_income)

# Calculate the percentage difference compared to black category
mean_income_by_color_exclude_other <- mean_income_by_color_exclude_other %>%
  mutate(percentage_difference = ((mean_income - reference_median) / reference_median) * 100) %>%
  
# Sort the data in ascending order of percentage difference
  arrange(percentage_difference)

mean_income_by_color_exclude_other

```

To compensate for this, we grouped our hair colors into broader
categories, such as combining different shades of blond and brown while
preserving other types as they were. We acknowledge the small count in
the "other" category. We will make further adjustments, setting the
stage for further analysis of income across these newly defined hair
color groups.

To begin analyzing this information, we calculate the average income for
the different hair color categories we are looking at. We create a bar
plot to visualize the mean income. As we mentioned earlier, the "other"
category was included in our dataset, and when it is graphed, it appears
misleading. We created a new plot and excluded the "other"
classification. We recalculated the mean income for the remaining
colors. Based on the information in our dataset, we saw that people with
blond hair, on average, have a higher mean income than those with black,
brown, or red hair.

## Combined Effects of Education and Hair on Income

```{r}
# Create a faceted scatterplot of income vs. education
p <- ggplot(data %>% filter(broad_hair_color != "other",  !is.na(income)), 
       aes(x = education, y = income)) +
  geom_point(aes(color = broad_hair_color)) +
  facet_wrap(~broad_hair_color) +
  labs(x = "Education Level", y = "Income", title = "Income vs. Education by Hair Color (Excluding 'Other' and Education < 95)") +
  scale_color_manual(values = hair_color_colors) +
  theme_bw() +
  guides(color = "none") +  # Remove the color legend
  scale_y_continuous(labels = scales::comma)

ggsave("Income_vs_Education_by_HairColor.png", plot = p)
p
```

In this analysis, we created a faceted scatterplot to explore the
relationship between income and education. We see how income and
education levels vary across different hair color groups. We exclude the
"other" hair color category to enhance clarity. While the scatter plot
reveals that individuals with blond hair, on average, tend to have lower
education completion than those with black or brown hair, they
consistently have a higher mean income. Additionally, it's worth noting
that higher education levels generally correlate with higher incomes
across all hair color categories. Further exploration of this data could
unveil more intriguing patterns and insights.

## Combined Effects of Gender and Hair on Income

```{r}

# Filter the data to exclude "other"
filtered_data <- data %>%
  filter(broad_hair_color != "other")

# Group the filtered data by broad_hair_color and gender, and calculate mean income
income_by_color_gender <- filtered_data %>%
  group_by(broad_hair_color, sex) %>%
  summarize(mean_income = mean(income, na.rm = TRUE))

# Define custom colors for gender
sex_color <- c(
  "male" = "blue", 
  "female" = "pink"
)

# Create a faceted bar plot for mean income by hair color and gender
p <- ggplot(income_by_color_gender, aes(x = broad_hair_color, y = mean_income, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Hair Color", y = "Mean Income", title = "Mean Income by Hair Color and Gender (Excluding 'Other')") +
  scale_fill_manual(values = sex_color) +  # Custom colors
  theme_minimal() +
  facet_wrap(~sex, scales = "free_y")  # Facet by gender with independent y-axes
ggsave("income_by_color_gender.png", plot = p)
p

# Group the data by broad_hair_color, sex, and calculate mean income
mean_income_by_color_gender <- data %>%
  group_by(broad_hair_color, sex) %>%
  summarize(mean_income = mean(income, na.rm = TRUE))


# Check the summary
mean_income_by_color_gender


# Calculate the mean income for each combination of broad_hair_color and sex
mean_income_by_color_gender <- data %>%
  group_by(broad_hair_color, sex) %>%
  summarize(mean_income = mean(income, na.rm = TRUE))

# Calculate the percentage difference
percentage_difference <- mean_income_by_color_gender %>%
  pivot_wider(names_from = sex, values_from = mean_income) %>%
  mutate(percentage_difference = ((male - female) / female) * 100)

# View the resulting data frame
percentage_difference


```

We created a faceted plot by filtering the hair color data to exclude
the "other" category in the dataset. We then grouped the variables
"broad_hair_color" and "sex" to calculate the mean income for each
resulting group. Upon reviewing the graph, it is evident that there are
gender-based income disparities. In most hair color categories, males
tend to have higher mean incomes than females. To provide a more
detailed understanding, we present the calculations in terms of
percentage variations in the subsequent tables, illustrating these
differences.

## Grouped Analysis of Summary Statistics

### Mean and Median Income by Gender

```{r}
data %>%
  group_by(sex) %>% summarise(Mean=mean(income, na.rm=T),
                              Median=median(income, na.rm=T))
```

Males seem to have higher average and median income compared with
females.

### Mean and Median Income by Hair

```{r}
data %>%
  group_by(hair) %>% summarise(Mean=mean(income, na.rm=T),
                              Median=median(income, na.rm=T))
```

Respondents with light blond color seem to have the highest average and
median income. Respondents with black color seem to have the lowest
average and median income.

### Mean and Median Income by Hair and Gender

```{r}
data %>%
  group_by(sex, hair) %>% summarise(Mean=mean(income, na.rm=T),
                              Median=median(income, na.rm=T))
```

### Proportion With Income More Than 200000 by Race

```{r}
data %>%
  group_by(sex) %>% summarise(Proportion=mean(income>200000, na.rm=T))
```

The proportion of income over \$200,000 is higher for males than
females.

### Proportion With Income More Than 200000 by Hair Color

```{r}
data %>%
  group_by(hair) %>% summarise(Proportion=mean(income>200000, na.rm=T))
```

The respondent with light blond color has the highest proportion of
income, more than \$200,000.

## Hypotheses for Further Analysis

### Hypothesis 1: Education and Income

We hypothesize that there is a positive relationship between education
level and income. In our analysis, we saw that individuals with higher
levels of education tended to have higher incomes. We expect that as
years of education increase, so does the average income.

### Hypothesis 2: Hair Color and Income

Based on our analysis of hair color and income, we hypothesize that
certain hair colors may be associated with higher incomes when compared
to others. Specifically, our data suggests that individuals with blond
hair tend to have a higher mean and median income than those with black,
brown, or red hair. This hypothesis may require further investigation to
test its statistical significance.

### Hypothesis 3: Gender-Based Income

We propose that gender-based income disparities exist in our dataset.
Specifically, we hypothesize that males, on average, have higher incomes
than females. Our hypothesis is drawn from our exploratory data
analysis, which reveals that the mean and median income for males is
consistently higher than that of females across various education
levels.
